
# Personalized logical models to study an interpret drug response

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\epigraph{"TBD."}{?}')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output(
  '>*TBD*  
  >?')
```

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\initial{T}he notion of modeling')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output('The notion of modeling.')
```

```{r 6_packages, echo=FALSE, warning=FALSE, message=FALSE}
invisible(lapply(X =  c("knitr", "tidyverse", "magrittr",  "ggplot2", "ggpubr", "patchwork", "cowplot", "kableExtra"),
       FUN = require,
       character.only = TRUE))

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  out.width = "90%",
  #fig.pos = "ht",
  fig.align = "center"
  )
```

```{block2, type='summarybox', echo = TRUE}

#### Publications {-}

This chapter extends the method presented in the previous chapter to investigate drug response with personalized logical models. The first application to cell lines of all cancer types was presented orally at ISMB2020 in Basel but is not published. The example concerning BRAF in melanomas and colorectal cancers is under revision and the corresponding pre-print is available as @beal2020personalized. Finally, the work on prostate cancer presented in a third section has also been submitted.

```

## One step further with drugs

### Methods

### Brute force

## A case study on BRAF in melanoma and colorectal cancers

## Limitations and perspectives illustrated by a prostate cancer study {#prostate-model}

A model is first of all an ambiguous object and a polysemous word. It therefore seems necessary to start with a semantic study. Among the many 


