
# Personalized logical models to study an interpret drug response

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\epigraph{"TBD."}{?}')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output(
  '>*TBD*  
  >?')
```

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\initial{T}he notion of modeling')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output('The notion of modeling.')
```

```{r 6_packages, echo=FALSE, warning=FALSE, message=FALSE}
invisible(lapply(X =  c("knitr", "tidyverse", "magrittr",  "ggplot2", "ggpubr", "patchwork", "cowplot", "kableExtra"),
       FUN = require,
       character.only = TRUE))

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  out.width = "90%",
  #fig.pos = "ht",
  fig.align = "center"
  )
```

## One step further with drugs

### Methods

### Brute force

## A case study on BRAF in melanoma and colorectal cancers

## Limitations and perspectives illustrated by a prostate cancer study {#prostate-model}

A model is first of all an ambiguous object and a polysemous word. It therefore seems necessary to start with a semantic study. Among the many 


