
# Some statistical evaluations about mechanistic models of cancer

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\epigraph{"Et l\'effet qui s\'en va nous decouvre les causes."}{Alfred de Musset (Poésies nouvelles, 1843)}')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output(
  '>*Et l\'effet qui s\'en va nous decouvre les causes.*<br/>
  >Alfred de Musset (Poésies nouvelles, 1843)')
```

```{r, include=knitr::is_latex_output(), echo=FALSE}
knitr::asis_output('\\initial{H}istorically, all mechanistic models of molecular networks, and logical models in particular, have been widely used to study response to treatments')
```

```{r, include=knitr::is_html_output(), echo=FALSE}
knitr::asis_output('Historically, all mechanistic models of molecular networks, and logical models in particular, have been widely used to study response to treatments ')
```

```{r 8_packages, echo=FALSE, warning=FALSE, message=FALSE}
invisible(lapply(X =  c("knitr", "tidyverse", "magrittr",  "ggplot2", "ggpubr", "patchwork", "cowplot"),
       FUN = require,
       character.only = TRUE))

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  out.width = "90%",
  #fig.pos = "ht",
  fig.align = "center"
  )
```
  


```{block2, type='summarybox', echo = TRUE}

#### Scientific content {-}

This chapter extends the method presented in the previous chapter to investigate drug response with personalized logical models. The first application to cell lines of all cancer types was presented orally at ISMB2020 in Basel but is not published.  

```

## One step further with drugs

One of the main clinical consequences of the underlying molecular complexity of cancers is the divergent response to treatment, even for a priori similar tumors. In the light of high-throughput sequencing data, the mechanisms governing these responses are somewhat better understood, for patients and especially for model organisms such as cell lines [@heiser2012subtype; garnett2012systematic]. But beyond a few simple cases, the diversity of response biomarkers once again calls for **holistic approaches** to unravel the underlying mechanisms.
